State-Trait Models

### 2024-06-06 examples/_3b_Examples_LinearMixedModels.qmd
updated output directory to the docs folder

```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered
#| cache: FALSE

@@

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, cbcl_scr_syn_external_r) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(cbcl_scr_syn_external_r) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = factor(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        cbcl_scr_syn_external_r = round(as.numeric(cbcl_scr_syn_external_r), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        height = cbcl_scr_syn_external_r
    )

df_long <- df_long %>%
  group_by(id) %>%
  filter(all(c("Baseline", "Year_1", "Year_2", "Year_3") %in% event)) %>%
  ungroup()

@@

### Descriptives
```{r}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered
#| cache: FALSE

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, cbcl_scr_syn_external_r) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(cbcl_scr_syn_external_r) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = factor(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        cbcl_scr_syn_external_r = round(as.numeric(cbcl_scr_syn_external_r), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        height = cbcl_scr_syn_external_r
    )

df_long <- df_long %>%
  group_by(id) %>%
  filter(all(c("Baseline", "Year_1", "Year_2", "Year_3") %in% event)) %>%
  ungroup()

@@

model <- lmer(height ~ event + (1 + event | id), data = df_long, REML = T)

# Model with random intercepts only
model_ri_only <- lmer(height ~ event + (1 | id), data = df_long, REML = T)
model_ri_only

# Model with random slopes only, without intercepts
model_rs_only <- lmer(height ~ event + (0 + event | id), data = df_long, REML = T)
model_rs_only

>>>>

```{r- BLOCK 1}
#| echo: FALSE # The content of this code block will not be shown when rendered
#| output: FALSE # Output from this code block will not be shown when rendered
#| cache: FALSE

@@

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, interview_age, anthroheightcalc) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(anthroheightcalc) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = factor(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        interview_age = as.numeric(interview_age),
        anthroheightcalc = round(as.numeric(anthroheightcalc), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        age = interview_age,
        height = anthroheightcalc
    )

@@

### Descriptives
```{r}

descriptives_table <- df_long %>%
    select(event, age, height) %>%
    mutate(event = factor(event)) %>%
    tbl_summary(
        by = event,
        missing = "no",
        label = list(height ~ "Height"),
        statistic = list(all_continuous() ~ "{mean} ({sd})", all_categorical() ~ "{p}%"),
    ) %>%
    modify_header(all_stat_cols() ~ "**{level}**<br>N = {n}") %>%
    bold_labels() %>%
    italicize_levels() %>%
    modify_spanning_header(all_stat_cols() ~ "**Assessment Wave**")
theme_gtsummary_compact()

descriptives_table

@@

df_long <- na.omit(df_long)

model <- lmer(height ~ age + (1 + age | id), data = df_long, REML = T)

# Model with random intercepts only
#model_ri_only <- lmer(height ~ age + (1 | id), data = df_long, REML = T)
#model_ri_only

# Model with random slopes only, without intercepts
#model_rs_only <- lmer(height ~ age + (0 + age | id), data = df_long, REML = T)
#model_rs_only

### 2024-06-06 examples/_queue/_11_Examples_MultivariateLatentGrowthCurveModels.qmd
updated output directory to the docs folder

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables
library(lavaan)       #
```

@@

# Specify the multivariate latent growth curve model
model <- '
  # Intercept and slope factors for variable 1 (height)
  i_var1 =~ 1*height_Baseline + 1*height_Year_1 + 1*height_Year_2
  s_var1 =~ 0*height_Baseline + 1*height_Year_1 + 2*height_Year_2

  # Intercept and slope factors for variable 2 (weight)
  i_var2 =~ 1*weight_Baseline + 1*weight_Year_1 + 1*weight_Year_2
  s_var2 =~ 0*weight_Baseline + 1*weight_Year_1 + 2*weight_Year_2
'

@@

@@

This plot illustrates xxxx.

>>>>

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables
library(lavaan)       #
library(patchwork)
```

@@

# Specify the multivariate latent growth curve model
model <- '
  # Intercept and slope factors for variable 1 (height)
  i_var1 =~ 1*height_Baseline + 1*height_Year_1 + 1*height_Year_2 + 1*height_Year_3
  s_var1 =~ 0*height_Baseline + 1*height_Year_1 + 2*height_Year_2 + 3*height_Year_3

  # Intercept and slope factors for variable 2 (weight)
  i_var2 =~ 1*weight_Baseline + 1*weight_Year_1 + 1*weight_Year_2 + 1*weight_Year_3
  s_var2 =~ 0*weight_Baseline + 1*weight_Year_1 + 2*weight_Year_2 + 3*weight_Year_3
'

@@

# Load necessary libraries
library(tidyverse)
library(patchwork)

# Calculate means for each time point for height and weight
mean_height <- df_wide %>%
  summarise(across(starts_with("height"), mean, na.rm = TRUE))

mean_weight <- df_wide %>%
  summarise(across(starts_with("weight"), mean, na.rm = TRUE))

# Reshape the data for plotting
mean_height_long <- pivot_longer(mean_height, cols = everything(), names_to = "Time", values_to = "Mean_Height")
mean_weight_long <- pivot_longer(mean_weight, cols = everything(), names_to = "Time", values_to = "Mean_Weight")

# Plot the mean trajectories for height
height_plot <- ggplot(mean_height_long, aes(x = Time, y = Mean_Height, group = 1)) +
  geom_line(color = "blue", size = 1.2) +
  geom_point(color = "blue") +
  labs(title = "Mean Growth Trajectory for Height", y = "Mean Height", x = "Time Point") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Plot the mean trajectories for weight
weight_plot <- ggplot(mean_weight_long, aes(x = Time, y = Mean_Weight, group = 1)) +
  geom_line(color = "red", size = 1.2) +
  geom_point(color = "red") +
  labs(title = "Mean Growth Trajectory for Weight", y = "Mean Weight", x = "Time Point") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Combine the plots side by side using patchwork
#combined_plot <- height_plot + weight_plot

# Display plots
print(height_plot)
print(weight_plot)

####

# Extract factor scores
factor_scores <- lavPredict(fit, type = "lv")

# Convert to a data frame
factor_scores_df <- as.data.frame(factor_scores)

# Rename the columns for clarity
colnames(factor_scores_df) <- c("Intercept_Height", "Slope_Height", "Intercept_Weight", "Slope_Weight")

# Summary of the model
summary(fit, fit.measures = TRUE, standardized = TRUE, rsquare = TRUE)

# Plot histograms of intercepts and slopes
ggplot(factor_scores_df, aes(x = Intercept_Height)) +
  geom_histogram(binwidth = 0.5, fill = "blue", alpha = 0.7) +
  labs(title = "Distribution of Intercepts for Height", x = "Intercept", y = "Frequency") +
  theme_minimal()

ggplot(factor_scores_df, aes(x = Slope_Height)) +
  geom_histogram(binwidth = 0.05, fill = "blue", alpha = 0.7) +
  labs(title = "Distribution of Slopes for Height", x = "Slope", y = "Frequency") +
  theme_minimal()

ggplot(factor_scores_df, aes(x = Intercept_Weight)) +
  geom_histogram(binwidth = 0.5, fill = "red", alpha = 0.7) +
  labs(title = "Distribution of Intercepts for Weight", x = "Intercept", y = "Frequency") +
  theme_minimal()

ggplot(factor_scores_df, aes(x = Slope_Weight)) +
  geom_histogram(binwidth = 0.05, fill = "red", alpha = 0.7) +
  labs(title = "Distribution of Slopes for Weight", x = "Slope", y = "Frequency") +
  theme_minimal()

#####

@@

These plots illustrate xxxx.

### 2024-06-06 examples/_queue/_13_Examples_GrowthMixtureModels.qmd
updated output directory to the docs folder

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, anthroheightcalc, anthroweightcalc) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(anthroheightcalc, anthroweightcalc) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = factor(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        anthroheightcalc = round(as.numeric(anthroheightcalc), 2),
        anthroweightcalc = round(as.numeric(anthroweightcalc), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        height = anthroheightcalc,
        weight = anthroweightcalc
    )

@@

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(lavaan)       #
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables

@@

>>>>

# Create and mutate the dataframe
df_long <- df %>%
    select(src_subject_id, eventname, anthroheightcalc, anthroweightcalc) %>%
    filter(eventname %in% c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1")) %>%
    drop_na(anthroheightcalc, anthroweightcalc) %>%
    arrange(src_subject_id, eventname) %>%
    mutate(
        src_subject_id = as.numeric(src_subject_id),
        eventname = factor(eventname, levels = c("baseline_year_1_arm_1", "1_year_follow_up_y_arm_1", "2_year_follow_up_y_arm_1", "3_year_follow_up_y_arm_1"), labels = c("Baseline", "Year_1", "Year_2", "Year_3")),
        anthroheightcalc = round(as.numeric(anthroheightcalc), 2),
        anthroweightcalc = round(as.numeric(anthroweightcalc), 2)
    ) %>%
    rename(
        id = src_subject_id,
        event = eventname,
        height = anthroheightcalc,
        weight = anthroweightcalc
    )

@@

## Preliminary Setup
### Load R libraries
```{r}
library(tidyverse)    # Collection of R packages for data science
library(lcmm)
library(lavaan)       #
library(rstatix)      # Pipe-friendly framework for basic statistical tests
library(DT)           # Rendering interactive data tables

@@

```{r}
set.seed(2002)
gmm1 <- hlme(height ~ event, subject = "id", random=~1, ng = 1, data =
df_long)

gmm2 <- gridsearch(rep = 100, maxiter = 10, minit = gmm1,
hlme(height ~ event, subject = "id_numeric", random=~1,
ng = 2, data = df_long, mixture = ~ event,
nwg=T))

gmm3 <- gridsearch(rep = 100, maxiter = 10, minit = gmm1,
hlme(height ~ event, subject = "id_numeric", random=~1,
ng = 3, data = df_long, mixture = ~ event,
nwg=T))

# make table with results for the 4 models:
summarytable(gmm1, gmm2, gmm3)
```

### 2024-06-06 examples/_queue/_14_Examples_StateTraitModels.qmd
updated output directory to the docs folder

# Define the model
model <- '
  # Trait component
  trait =~ 1 * anthroheightcalc_0 + 1 * anthroheightcalc_1 + 1 * anthroheightcalc_2 + 1 * anthroheightcalc_3

  # State component
  state =~ anthroheightcalc_0 + anthroheightcalc_1 + anthroheightcalc_2 + anthroheightcalc_3

  # Residual variances
  anthroheightcalc_0 ~~ anthroheightcalc_0
  anthroheightcalc_1 ~~ anthroheightcalc_1
  anthroheightcalc_2 ~~ anthroheightcalc_2
  anthroheightcalc_3 ~~ anthroheightcalc_3
'

>>>>

# Define the model
model <- '
  # Trait component
  trait =~ 1 * height_Baseline + 1 * height_Year_1 + 1 * height_Year_2 + 1 * height_Year_3

  # State component
  state =~ height_Baseline + height_Year_1 + height_Year_2 + height_Year_3

  # Residual variances
  height_Baseline ~~ height_Baseline
  height_Year_1 ~~ height_Year_1
  height_Year_2 ~~ height_Year_2
  height_Year_3 ~~ height_Year_3
'
